---
title: "Learn more about EARTHQUAKES LM"
author: "Sujoy Basu"
date: "April 24, 2016"
output: html_document
---

This application makes it easy to take a dataset in CSV format and build a linear model using the lm command in R. The data scientist can view relevant analysis of the data on the single page of the application, and also search through the dataset . The current version of this application uses the earthquakes dataset installed with the application. Future versions will allow the dataset to be uploadeed. 

To use this application, you have to make selections in the left panel:

1. Select the column that will be the response column from the dropdown menu. The default is magnitude from the earthquakes dataset. This column will be expressed as a linear function of the variables selected in the next step. 

2. Select the variables, that are also referred to as the terms for the linear predictor. Multiple checkboxes can be selected. The defaults are depth and NST columns of the earthquakes dataset.

The output of this application is reactive. As a result of calculations done by the shiny server, the display is updated. For the default selections, add longitude ('lon') to the terms of the linear predictor, and observe how the top four sections of the output are updated. 

The output is divided into the following five sections:

A. The top section displays the formula for the linear model.

```{r, echo=FALSE}
print("Formula (frmlObj): ")
      frmlObj <- as.formula(paste('Magnitude'," ~ ",paste(c('Depth', 'NST'),collapse="+")))
      print(frmlObj, showEnv=FALSE)
```

B. The second section displays multiple plots that are obtained from the linear model, and allow the data scientist to evaluate the model based on the residuals, as shown in this example:

```{r, echo=FALSE}
df <- read.csv("earthquakes.csv", header = TRUE, quote="\"", stringsAsFactors= FALSE, strip.white = TRUE)
frmlObj <- as.formula(paste('Magnitude'," ~ ",paste(c('Depth', 'NST'),collapse="+")))
lmObject <- lm(frmlObj,data=df,model=TRUE)
par(mfrow = c(2, 2), oma = c(0, 0, 2, 0))
plot(lmObject, id.n = 5, labels.id = NULL) 
```

C. The next section summarizes the linear model by invoking the summary method of the lm class. Whether the first metric you look at to evaluate the model is the Adjusted R-squared metric, or the distribution of the residuals, or something else, you are likely to find it in this section:

```{r, echo=FALSE}
summary.lm(lmObject, correlation=TRUE, digits=5, signif.stars=TRUE)
```

D. The following section summarizes the ANOVA analysis:

```{r, echo=FALSE}
anova(lmObject)
```

E. The fifth section, which is in a separate tab, displays the content of the CSV file, and allows data points to be searched.

